### 2023-04-01 R/Paper_plots_lenticel_damage.Rmd
before last revision

---
title: "Paper_Cartama_lenticel_damage"
author: "dru"
date: "2/3/2022"
output:
  pdf_document: default
  html_document: default
---

## Preparing the environment

Loading required Packages

```{r Packages, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(gridExtra)
library(readxl)
library(ggpubr)
library(tibble)
library(ggplot2)
library(extrafont)
loadfonts(device = "win")
library(ggsignif)
library(ggpubr)

```

## Loading and inspecting the data

Loading and inspecting the file 'Lenticel_Damage_Summary_All_Harvest.xlsx'. This file has two-years of lenticel damage measurements recorded in two avocado farms across four harvest.

```{r echo=TRUE, message=FALSE, warning=FALSE}
Lenticel_file <- read_excel("C:/Users/Dru/Desktop/Cartama_paper/files/Lenticel_Damage_Summary_All_Harvest.xlsx")
glimpse(Lenticel_file)
```
Checking for NA, miss speels and anomalies in the data:

```{r inspecting}

table(Lenticel_file$Cosecha)
table(Lenticel_file$UP)

f <-subset(Lenticel_file,Lenticel_file$UP=="Escondida")
f1 <- subset(Lenticel_file,Lenticel_file$UP=="Sinai")
table(f$Lote)
table(f$Arbol)
table(f1$Lote)
table(f1$Arbol)
summary(Lenticel_file$Severidad_0)
summary(Lenticel_file$Severidad_1)
summary(Lenticel_file$Incidencia_0)
summary(Lenticel_file$Incidencia_1)

```
Problems:

* Some tree numbers are shared between La Escondida y El Sinai, and this redundancy can result in problems during the analysis.
* Lenticel damage columns have several NA. Fruits having NA the severidad_0 or incidence_0 were removed from all the analyses, and those having NA in severidad_1 or incidence_1 only from the analysis evaluating disease progresion

## Cleaning and preparing the data
### cleaning actions:

* The columns cosecha, Año, Severidad_pedunculo,Fruta and the last three empty columns were removed as they are not used during the analysis
* The 'p' simbol in the column Cosecha is changed for 'm' to especify the main harvest.
* Generated a column Tree which has the name of the tree indicated by the farm and tree number (i.e., Sinai15) to avoid problems during the analysis
* Removed the Arbol and Fruta columns

```{r cleaning}
Lenticel_file$...13<-NULL
Lenticel_file$...14<-NULL
Lenticel_file$...15<-NULL
Lenticel_file$Año<-NULL
Lenticel_file$cosecha<-NULL
Lenticel_file$Fruta<-NULL
Lenticel_file$Severidad_pedunculo<-NULL
Lenticel_file[Lenticel_file=="2019_p"]<-"2019_m"
Lenticel_file[Lenticel_file=="2020_p"]<-"2020_m"
Lenticel_file$Tree<-paste(Lenticel_file$UP,Lenticel_file$Arbol,sep="")
Lenticel_file$Arbol<-NULL

df <- Lenticel_file %>% drop_na(Severidad_0 , Incidencia_0)
colnames(df)[which(names(df) == "Cosecha")] <- "Harvest"
colnames(df)[which(names(df) == "UP")] <- "Farm"
colnames(df)[which(names(df) == "Lote")] <- "Plot"
colnames(df)[which(names(df) == "Severidad_0")] <- "Sev_0dph"
colnames(df)[which(names(df) == "Severidad_1")] <- "Sev_21dph"
colnames(df)[which(names(df) == "Incidencia_0")] <- "Inc_0dph"
colnames(df)[which(names(df) == "Incidencia_1")] <- "Inc_21dph"
df$Inc_increase <- df$Inc_21dph-df$Inc_0dph
df$Sev_increase <- df$Sev_21dph-df$Sev_0dph

glimpse(df)
summary(df$Sev_0dph)

```
## Processing and plottin the lenticel damage data at 0 dph for the farms and harvest

* The mean of the severity and incidence of the lenticel damage was calculated grouping by Plot and Harvest

```{r df_for_barplot}

df_mean <- tapply(df$Sev_0dph , list(df$Plot , df$Harvest) , mean)
sev <- df_mean[1 , ]
df_mean1 <- tapply(df$Inc_0dph , list(df$Plot , df$Harvest) , mean)
inc <- df_mean1[1 , ]

for (i in 2:13){
  sev<-c(sev , df_mean[i,])
  inc <- c(inc , df_mean1[i,])
  }
mat <- data.frame(rep(c("2019_m" , "2020_m" , "2020_t" , "2021_t") , 13) , c(rep("Escondida" , 20) , rep("Sinai" , 32)) , sev , inc)

colnames(mat) <- c("Harvest" , "Farm" , "Severity" , "Incidence")

print(head(mat))

```
* The barplot showing the severity of lenticel damage for the farms across harvest Harvest. The bars and error bars represent the severity mean and standard deviation observed in the farms for each harvest at 0 dph, and the points represent the severity mean for the plots in the farm.

```{r severity plot}

p <-  ggplot(mat, aes(x = factor(Harvest,levels = c("2019_m","2020_t","2020_m","2021_t")), y = Severity)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(), fill=rep(c("gray90","gray70"),4),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar", color="black", width=0.2) 
p <- p + geom_jitter()+ facet_wrap(~Farm) + theme_classic() 
p <- p +  xlab("") + ylab("Necrotic Area (%)")
p <- p + theme(axis.title =element_text(size=14,family = 'serif'))
p <- p +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,family = 'serif'),
              axis.text.y = element_text(face="bold",size=12, angle=0,family = 'serif'))
p <- p + theme(strip.text.x = element_text(size = 14,family = 'serif'))
p

```

* The barplot showing the Incidence of lenticel damage for the farms across harvest Harvest. The bars and error bars represent the severity mean and standard deviation observed in the farms for each harvest at 0 dph, and the points represent the incidence mean for the plots in the farm.

```{r Incidence plot}

p1 <-  ggplot(mat, aes(x = factor(Harvest,levels = c("2019_m","2020_t","2020_m","2021_t")), y = Incidence)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(), fill=rep(c("gray90","gray70"),4),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar", color="black", width=0.2) 
p1 <- p1 + geom_jitter()+ facet_wrap(~Farm) + theme_classic() 
p1 <- p1 +  xlab("") + ylab("Number of necrotic spots")
p1 <- p1 + theme(axis.title =element_text(size = 14 , family = 'serif'))
p1 <- p1 +  theme(axis.text.x = element_text(face="bold", size = 12, angle = 0 , family = 'serif'),
              axis.text.y = element_text(face = "bold" , size = 12 , angle = 0 , family = 'serif'))
p1 <- p1 + theme(strip.text.x = element_text(size = 14 , family = 'serif'))
p1

```

* Fig 1.* Lenticel damage for avocado cv. Hass fruits collected from the La Escondida and El Sinai farms between 2019 and 2021 at harvest time (0 days post-harvest). Bars and errorbars represent mean and standard deviation for the severity (A) and incidence (B) of the damage observed in the farms for each harvest at 0 dph. The points show the severity and incidence means for the fruits collected from the different plots, and the letter after the year denotes main (m) and traviesa (t) harvest.

```{r severity and incidence plot}

 plot<-grid.arrange(p + labs(tag = 'A'),p1 + labs(tag = 'B'), ncol=1,nrow=2)
ggsave("C:/Users/Dru/Desktop/Cartama_paper/Figure1_new.jpg",plot=plot,dpi = 300,width = 7,height = 7,units = c("in"))

```

## Processing and plottin the lenticel damage increase between 0 dph and 21 dpc

* Note: The analisis omitted fruits with no record of lenticel damage (NA) at 21 dph
* were omitted from  was calculated grouping by Plot and Harvest

```{r df_for_barplot plots}

df1 <- na.omit(df)

sev <- c(df1$Sev_0dph , df1$Sev_21dph)
inc <- c(df1$Inc_0dph , df1$Inc_21dph)

###stack plot
#sev <- c(df1$Sev_0dph , df1$Sev_increase)
#inc <- c(df1$Inc_0dph , df1$Inc_increase)
farm <- c(df1$Farm , df1$Farm)
time <-factor(c(rep("0dpc",length(sev)/2) , rep("21dpc",length(sev)/2)))
harvest <- c(df1$Harvest , df1$Harvest)

mat <- data.frame(Farm = farm, Time = time , Harvest = harvest, Severity = sev , Incidence = inc)

mat_mean<-tapply(mat$Severity , list(mat$Time,mat$Harvest,mat$Farm),mean)
mat_mean1<-tapply(mat$Incidence , list(mat$Time,mat$Harvest,mat$Farm),mean)

mat1<-data.frame(Farm = c(rep("Escondida" , 8) , rep("Sinai" , 8)), Time = rep(c(rep("0dpc",4) , rep("21dpc",4)) , 2) ,
                 Harvest = rep(c("2019_m" , "2020_m" , "2020_t" , "2021_t") , 4), 
                 Severity = c(mat_mean[1,,1] , mat_mean[2,,1] , mat_mean[1,,2] , mat_mean[2,,2]),
                 Incidence = c(mat_mean1[1,,1] , mat_mean1[2,,1] , mat_mean1[1,,2] , mat_mean1[2,,2]) )

head(mat1)

```
* The barplot showing the increase in the severity of lenticel damage during the cold storage. The bars and error bars represent the severity mean and standard deviation observed in the farms at the two evaluations (0 dph and 21 dph), and the points represent the severity mean for the harvest in the farm.

```{r severity in time}
p <-  ggplot(mat1, aes(x = factor(Time), y = Severity)) + 
      geom_bar(stat="summary", color="black", positio=position_dodge(), fill= rep(c("gray99","gray95") ,2), fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar", color="black", width=0.2) 
p <- p + geom_jitter(aes(colour = factor(Harvest , levels = c("2019_m" , "2020_t" , "2020_m" , "2021_t"))) , size = 2)
p <- p + facet_wrap(~ Farm) + theme_classic() 
p <- p +  xlab("") + ylab("Necrotic Area (%)")
p <- p + theme(axis.title =element_text(size=14,family = 'serif'))
p <- p +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,family = 'serif'),
              axis.text.y = element_text(face="bold",size=12, angle=0,family = 'serif'))
p <- p + theme(strip.text.x = element_text(size = 14,family = 'serif'))
p <- p + scale_color_manual(name = "Harvest",  values =  c("gray88","gray70","gray47","gray0"))
p <- p + theme(legend.text = element_text(size=12, angle=0,family = 'serif'),
                legend.title = element_text(face="bold",size=12, angle=0,family = 'serif'))
p

```
* The barplot showing the increase in the Incidence of lenticel damage during the cold storage. The bars and error bars represent the incidence mean and standard deviation observed in the farms at the two evaluations (0 dph and 21 dph), and the points represent the severity mean for the harvest in the farm.

```{r incidence in time}
p1 <-  ggplot(mat1, aes(x = factor(Time), y = Incidence)) + 
      geom_bar(stat="summary", color="black", positio=position_dodge(), fill= rep(c("gray99","gray95") ,2), fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar", color="black", width=0.2) 
p1 <- p1 + geom_jitter(aes(colour = factor(Harvest , levels = c("2019_m" , "2020_t" , "2020_m" , "2021_t"))) , size = 2)
p1 <- p1 + facet_wrap(~ Farm) + theme_classic() 
p1 <- p1 +  xlab("") + ylab("Number of necrotic spots")
p1 <- p1 + theme(axis.title =element_text(size=14,family = 'serif'))
p1 <- p1 +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,family = 'serif'),
              axis.text.y = element_text(face="bold",size=12, angle=0,family = 'serif'))
p1 <- p1 + theme(strip.text.x = element_text(size = 14,family = 'serif'))
p1 <- p1 + scale_color_manual(name = "Harvest", values =  c("gray88","gray70","gray47","gray0"))
p1 <- p1 + theme(legend.text = element_text(size=12, angle=0,family = 'serif'),
                legend.title = element_text(face="bold",size=12, angle=0,family = 'serif'))

```
*Fig 2.* Lenticel damage for avocado cv. Hass fruits collected from the La Escondida and El Sinai farms between 2019 and 2021 at harvest time (0 days post-harvest, 0 dph) and after 21 days of storage at 6C (21 days post-harvest, 21 dph). Shown are the means and standard deviations for the severity (A) and incidence (B) of the damage for the two evaluatin points (0 dph and 21 dph). The points show the severity and incidence means for the fruits collected in the four harvest, and the letter after the year denotes main (m) and traviesa (t) harvest.

```{r severity and incidence in time plot}

 plot<-grid.arrange(p + labs(tag = 'A'),p1 + labs(tag = 'B'), ncol=1,nrow=2)
ggsave("C:/Users/Dru/Desktop/Cartama_paper/Figure2_new.jpg",plot=plot,dpi = 300,width = 7,height = 7,units = c("in"))

```
```{r df_for_barplot plots}

df1 <- na.omit(df)

sev <- c(df1$Sev_0dph , df1$Sev_21dph)
inc <- c(df1$Inc_0dph , df1$Inc_21dph)

farm <- c(df1$Farm , df1$Farm)
time <-factor(c(rep("0dph",length(sev)/2) , rep("21dph",length(sev)/2)))
harvest <- c(df1$Harvest , df1$Harvest)

mat <- data.frame(Farm = farm, Time = time , Harvest = harvest, Severity = sev , Incidence = inc)

mat_mean<-tapply(mat$Severity , list(mat$Time,mat$Harvest,mat$Farm),mean)
mat_mean1<-tapply(mat$Incidence , list(mat$Time,mat$Harvest,mat$Farm),mean)
mat_se<-tapply(mat$Severity , list(mat$Time,mat$Harvest,mat$Farm),sd)/sqrt(table(mat$Time,mat$Harvest,mat$Farm))
mat_se1<-tapply(mat$Incidence , list(mat$Time,mat$Harvest,mat$Farm),sd)/sqrt(table(mat$Time,mat$Harvest,mat$Farm))
n <- table(mat$Time,mat$Harvest,mat$Farm)

mat1<-data.frame(Farm = c(rep("Escondida" , 8) , rep("Sinai" , 8)), Time = rep(c(rep("0dph",4) , rep("21dph",4)) , 2) ,
                 Harvest = rep(c("2019_m" , "2020_m" , "2020_t" , "2021_t") , 4), 
                 Severity = c(mat_mean[1,,1] , mat_mean[2,,1] , mat_mean[1,,2] , mat_mean[2,,2]),
                 Severity_se = c(mat_se[1,,1] , mat_se[2,,1] , mat_se[1,,2] , mat_se[2,,2]),
                 Incidence = c(mat_mean1[1,,1] , mat_mean1[2,,1] , mat_mean1[1,,2] , mat_mean1[2,,2]),
                 Incidence_se = c(mat_se1[1,,1] , mat_se1[2,,1] , mat_se1[1,,2] , mat_se1[2,,2]))

head(mat1)

```
* The barplot showing the increase in the severity of lenticel damage during the cold storage. The bars and error bars represent the severity mean and standard deviation observed in the farms at the two evaluations (0 dph and 21 dph), and the points represent the severity mean for the harvest in the farm.

```{r severity in time}

data_hline <- data.frame(Farm = c(rep("Escondida",4),rep( "Sinai",4)),
                         x = c(0.7,1.7,2.7,3.7,0.7,1.7,2.7,3.7), y = c(1.2, 2.4,0.9,4.7,2.8,4.6,5.4,3.6),xend = c(1.3,2.3,3.3,4.3,1.3,2.3,3.3,4.3), yend = c(1.2, 2.4,0.9,4.7,2.8,4.6,5.4,3.6), label = rep("*",8))

p <-  ggplot(mat1, aes(x = factor(Harvest, levels = c("2019_m" , "2020_t" , "2020_m" , "2021_t")), y = Severity, fill = factor(Time))) + geom_bar(stat="identity", positio=position_dodge())     
p <- p + scale_fill_manual(values=c("gray90", "gray80"))
p <- p + geom_errorbar(aes(ymin=Severity-Severity_se, ymax=Severity+Severity_se), width=.2,position=position_dodge(.9)) 
p <- p + facet_grid(~ Farm) + theme_classic() 
p <- p +  xlab("") + ylab("Necrotic Area (%)") + labs(fill='Days postharvest') 
p <- p + theme(axis.title =element_text(size=14,family = 'serif'))
p <- p +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,family = 'serif'),
              axis.text.y = element_text(face="bold",size=12, angle=0,family = 'serif'))
p <- p + theme(strip.text.x = element_text(size = 14,family = 'serif'))
p <- p + theme(legend.text = element_text(size=12, angle=0,family = 'serif'),
                legend.title = element_text(face="bold",size=12, angle=0,family = 'serif'))

p <- p + geom_segment(data = data_hline,aes(x = x,y = y,xend = xend, yend = yend),inherit.aes = FALSE) 
p <- p +geom_text(data = data_hline,aes(label = label,x = x+0.3,y = y+0.2),inherit.aes = FALSE)

```
* The barplot showing the increase in the Incidence of lenticel damage during the cold storage. The bars and error bars represent the incidence mean and standard deviation observed in the farms at the two evaluations (0 dph and 21 dph), and the points represent the severity mean for the harvest in the farm.

```{r incidence in time}

p1 <-  ggplot(mat1, aes(x = factor(Harvest, levels = c("2019_m" , "2020_t" , "2020_m" , "2021_t")), y = Incidence, fill = factor(Time))) + geom_bar(stat="identity", positio=position_dodge())     
p1 <- p1 + scale_fill_manual(values=c("gray90", "gray80"))
p1 <- p1 + geom_errorbar(aes(ymin=Incidence-Incidence_se, ymax=Incidence+Incidence_se), width=.2,position=position_dodge(.9)) 
p1 <- p1 + facet_wrap(~ Farm) + theme_classic() 
p1 <- p1 +  xlab("Harvest") + ylab("Number of necrotic spots") + labs(fill='Days postharvest') 
p1 <- p1 + theme(axis.title =element_text(size=14,family = 'serif'))
p1 <- p1 +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,family = 'serif'),
              axis.text.y = element_text(face="bold",size=12, angle=0,family = 'serif'))
p1 <- p1 + theme(strip.text.x = element_text(size = 14,family = 'serif'))
p1 <- p1 + scale_color_manual(name = "Harvest", values =  c("gray88","gray70","gray47","gray0"))
p1 <- p1 + theme(legend.text = element_text(size=12, angle=0,family = 'serif'),
                legend.title = element_text(face="bold",size=12, angle=0,family = 'serif'))

data_hline <- data.frame(Farm = c(rep("Escondida",4),rep( "Sinai",4)),
                         x = c(0.7,1.7,2.7,3.7,0.7,1.7,2.7,3.7), y = c(88, 165,70,203,238,217,256,187),xend = c(1.3,2.3,3.3,4.3,1.3,2.3,3.3,4.3), yend = c(88, 165,70,203,238,217,256,187), label = rep("*",8))

p1 <- p1 + geom_segment(data = data_hline,aes(x = x,y = y,xend = xend, yend = yend),inherit.aes = FALSE) 
p1 <- p1 + geom_text(data = data_hline,aes(label = label,x = x+0.3,y = y+5),inherit.aes = FALSE)
p1

```
*Fig 2.* Lenticel damage for avocado cv. Hass fruits collected from the La Escondida and El Sinai farms between 2019 and 2021 at harvest time (0 days post-harvest, 0 dph) and after 21 days of storage at 6C (21 days post-harvest, 21 dph). Shown are the means and standard errors for the severity (A) and incidence (B) of the damage for the two evaluatin points (0 dph and 21 dph). The * denotes statistical differences at a 95% confidence level according to the  generalized linear mixed models, and the letter after the year denotes main (m) and traviesa (t) harvest.

```{r severity and incidence in time plot}

 plot<-grid.arrange(p + labs(tag = 'A'),p1 + labs(tag = 'B'), ncol=1,nrow=2)
ggsave("C:/Users/Dru/Desktop/Cartama_paper/Figure1_new.jpg",plot=plot,dpi = 300,width = 7,height = 7,units = c("in"))

```