### 2023-04-01 R/Paper_plots_microbial_ecology.Rmd

---
title: "Paper_Cartama_Microbial_ecology"
author: "dru"
date: "2/16/2022"
output:
  pdf_document: default
  html_document: default
---

## Preparing the environment

Loading required Packages

```{r Packages, message=FALSE, warning=FALSE}

library(tidyverse)
library(gridExtra)
library(tibble)
library(ggplot2)
library(extrafont)
loadfonts(device = "win")
library(phyloseq)
library(scales)   
library(ggpubr)
library(readxl)
library(vegan)
library(biomformat)
library(patchwork)
library(car)
library(ggimage)
library(ggtree)
library(TDbook)
library(aplot)
library(lemon)
library(grid, lib.loc = "C:/Program Files/R/R-4.2.0/library")
setwd("C:/Users/Dru/Desktop/Cartama_paper/microbial-ecology-avocado/Resultados_all")
```

## Lenticel damage the fruits

Generating the dataframe containing the lenticel damage data for the differentialy diseased fruits used inthe microbial ecology annalisis

```{r data}

md <- data.frame("Severity"=c(.7 , .84 , 1.66 , 1.66 ,0.62 , 2.02 , 0.085 , 0.025 , 0.08 ,
                              0.085 ,0.095 , 0.285 , 23.4 , 17.97 , 15.99 , 19.13 , 19.01,
                              12.92 , 1.73 , 0.18 , 0.81 , 1.15 , 0.47 , 0.09),
                 "Incidence"=c(50 , 84 , 89 , 133 , 84 , 158 , 11 , 5 , 12 ,30 , 14 , 31 ,
                               525 , 488 , 347 , 269 , 358 , 515 , 128 , 62 ,119 , 94 ,
                               108 , 68),
                 "Farm"= c(rep("Escondida" , 12) , rep("Siani" , 12)),
                 "Damage"= c(rep(c(rep("Severe" , 6) , rep("Mild" , 6)),2)))

head(md)

```
Ploting the severity data:

```{r severity plot}

p <-  ggplot(md, aes(x = Damage, y = Severity)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(),
               fill=rep(c("gray90","gray70"),2),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar", color="black", width=0.2)
p <- p + geom_jitter()+ facet_wrap(~Farm) + theme_classic() 
p <- p +  xlab("") + ylab("Necrotic Area (%)")+ylim(c(0, max(md$Severity)*1.2))
p <- p + theme(axis.title =element_text(size=14,family = 'serif'))
p <- p +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,family = 'serif'),
              axis.text.y = element_text(face="bold",size=12, angle=0,family = 'serif'))
p <- p + theme(strip.text.x = element_text(size = 14,family = 'serif'))

p <- p +     geom_signif(test="t.test", comparisons = list(c("Severe", "Mild")), map_signif_level = T)

```
Ploting the incidence data:

```{r incidence plot}

p1 <-  ggplot(md, aes(x = Damage, y = Incidence)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(),
               fill=rep(c("gray90","gray70"),2),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar", color="black", width=0.2) 
p1 <- p1 + geom_jitter()+ facet_wrap(~Farm) + theme_classic() 
p1 <- p1 +  xlab("") + ylab("Necrotic spots")+ylim(c(0, max(md$Incidence)*1.2))
p1 <- p1 + theme(axis.title =element_text(size=14,family = 'serif'))
p1 <- p1 +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,family = 'serif'),
              axis.text.y = element_text(face="bold",size=12, angle=0,family = 'serif'))
p1 <- p1 + theme(strip.text.x = element_text(size = 14,family = 'serif'))
p1 <- p1 +     geom_signif(test="t.test", comparisons = list(c("Severe", "Mild")), map_signif_level = T)
```
*Supplementary figure 1.* Lenticel damage for avocado cv. Hass fruits used for the microbial-community analysis. Fruits were collected from the La Escondida and El Sinai farms during the traviesa harvest of 2020. Shown are the means and standard deviations for the severity (A) and incidence (B) of the damage for the fruits with deferent levels of affection. The points show the severity and incidence for the fruits and the asterisk denote statistic differences at the 99.0% () and 99.9% () confidence level according to the t test (n: 6).

```{r severity and incidence plot}

 plot<-grid.arrange(p + labs(tag = 'A'),p1 + labs(tag = 'B'), ncol=1,nrow=2)
ggsave("C:/Users/Dru/Desktop/Cartama_paper/SupplementaryFig1.jpg",plot=plot,dpi = 300,width = 7,height = 7,units = c("in"))
```
## Rarefraction curve plots

```{r}

#Curvas de rarefacci?n antes de normalizacion
biomfile_before <- import_biom("ASVs_table6_dada2R.biom")
biomfile_after <- import_biom("ASVs_table9_dada2R.biom")
head(otu_table(biomfile))
head(tax_table(biomfile))

otu.table_before <- otu_table(biomfile_before)
t.OTU.table_before <- t(otu.table_before)

otu.table_after <- otu_table(biomfile_after)
t.OTU.table_after <- t(otu.table_after)
#head(otu_table(biomfile_before))
#head(tax_table(biomfile))

col <- c("black", "darkred", "forestgreen", "orange", "blue", "yellow", "hotpink")
lty <- c(rep("solid",6),rep( "dashed",6), rep( "longdash", 6), rep("dotdash",6))
#par(mar=c(5,4,2,2))
#par(mgp=c(2.3,0.5,0))
jpeg("rarre.jpg", units="in", width=7, height=6, res=300)
par(mfrow=c(1,2))
p<-rarecurve(t.OTU.table_before, col="black", step = 1000, cex = 0.8, lwd=1, label = FALSE,xlab = "Sample size", ylab = "Specie Number", ylim=c(0, 680),family="serif",cex.lab=1.4, cex.axis=1.2,lty=lty)

text(-14800, 760, "A", family = "serif", xpd=NA, cex = 1.2 )
p1<-rarecurve(t.OTU.table_after, col="black", step = 1000, cex = 0.8, lwd=1, label = FALSE,xlab = "Sample size", ylab = "Specie Number", ylim=c(0, 680),family="serif",cex.lab=1.4, cex.axis=1.2,lty=lty)

text(-6000, 760, "B", family = "serif", xpd=NA, cex = 1.2 )
dev.off()

#The rarefaction curves are evaluated using the interval of step sample sizes
#Graphical Parameters: https://www.statmethods.net/advgraphs/parameters.html

```

```{r severity and incidence plot}

jpeg("rarre.jpg",, units="in", width=5, height=5, res=300)
dev.off()
```

## Alpha diversity plot

Loading the dataframe containing the alpha diversity metrics for the for the differentialy diseased fruits

```{r alfa df}
path<-"C:/Users/Dru/Desktop/Cartama_paper/microbial-ecology-avocado/Resultados_all/Alfa/"
Alfa <- read_excel(paste(path,"Alfa.xlsx",sep=""))
head(Alfa)
```
```{r alfa analisis}
c <- subset(Alfa, Alfa$Farm == "Sinai")$Chao
s <- subset(Alfa, Alfa$Farm == "Sinai")$Shannon_entropy
r <- subset(Alfa, Alfa$Farm == "Sinai")$Riqueza
f <- subset(Alfa, Alfa$Farm == "Sinai")$Faith_pd
e <- subset(Alfa, Alfa$Farm == "Sinai")$Evenness

damage <-subset(Alfa, Alfa$Farm == "Escondida")$Damage_level

#names(shannon)
#Check ANOVA assumptions
#Homogeneity of variance
leveneTest(c ~ damage)  #p-value > 0.05 
#Shapiro-Wilk test
shapiro.test(c) #p-value > 0.05 is normal
#Independence
durbinWatsonTest(lm(c ~ damage)) #p-value > 0.05 they are independent

t.test(r~damage)
kruskal.test(s~damage)
summary(glm(c~damage, family = quasipoisson()))
plot(glm(c~damage, family = poisson()))
plot(lm(c~damage))
```
Ploting the alpha_diversity metrics

```{r alpha-diversity plots}

#### Chao

c <-  ggplot(Alfa, aes(x = Damage_level, y = Chao)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(),
               fill=rep(c("gray90","gray70"),2),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar",
                   color="black", width=0.2) 
c <- c + geom_jitter()+ facet_wrap(~Farm)+ theme_classic() 
c <- c +  xlab("") + ylab("Chao")+ylim(c(0, max(Alfa$Chao)*1.1))
c <- c + theme(axis.title =element_text(size=14,family = 'serif'))
c <- c +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,
                                             family ='serif'),
                  axis.text.y = element_text(face="bold",size=12, angle=0,family =
                                               'serif'))
c <- c + theme(strip.text.x = element_text(size = 14,family = 'serif'))
c <- c + geom_signif(test="kruskal.test", comparisons = list(c("Mild", "Severe")), map_signif_level = T)

#### Shannon_entropy

s <-  ggplot(Alfa, aes(x = Damage_level, y = Shannon_entropy)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(),
               fill=rep(c("gray90","gray70"),2),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar",
                   color="black", width=0.2) 
s <- s + geom_jitter()+ facet_wrap(~Farm)+ theme_classic() 
s <- s +  xlab("") + ylab("Shannon entropy")+ylim(c(0, max(Alfa$Shannon_entropy)*1.1))
s <- s + theme(axis.title =element_text(size=14,family = 'serif'))
s <- s +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,
                                             family ='serif'),
                  axis.text.y = element_text(face="bold",size=12, angle=0,family =
                                               'serif'))
s <- s + theme(strip.text.x = element_text(size = 14,family = 'serif'))
s <- s + geom_signif(test="kruskal.test", comparisons = list(c("Mild", "Severe")), map_signif_level = T)

#### Riqueza

r <-  ggplot(Alfa, aes(x = Damage_level, y = Riqueza)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(),
               fill=rep(c("gray90","gray70"),2),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar",
                   color="black", width=0.2) 
r <- r + geom_jitter()+ facet_wrap(~Farm)+ theme_classic() 
r <- r +  xlab("") + ylab("Richness")+ylim(c(0, max(Alfa$Riqueza)*1.1))
r <- r + theme(axis.title =element_text(size=14,family = 'serif'))
r <- r +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,
                                             family ='serif'),
                  axis.text.y = element_text(face="bold",size=12, angle=0,family =
                                               'serif'))
r <- r + theme(strip.text.x = element_text(size = 14,family = 'serif'))
r <- r + geom_signif(test="kruskal.test", comparisons = list(c("Mild", "Severe")), map_signif_level = T)

#### Faith_pd

f <-  ggplot(Alfa, aes(x = Damage_level, y = Faith_pd)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(),
               fill=rep(c("gray90","gray70"),2),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar",
                   color="black", width=0.2) 
f <- f + geom_jitter()+ facet_wrap(~Farm)+ theme_classic() 
f <- f +  xlab("") + ylab("Faith pd")+ylim(c(0, max(Alfa$Faith_pd)*1.1))
f <- f + theme(axis.title =element_text(size=14,family = 'serif'))
f <- f +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,
                                             family ='serif'),
                  axis.text.y = element_text(face="bold",size=12, angle=0,family =
                                               'serif'))
f <- f + theme(strip.text.x = element_text(size = 14,family = 'serif'))
f <- f + geom_signif(test="kruskal.test", comparisons = list(c("Mild", "Severe")), map_signif_level = T)

###evenness
e <-  ggplot(Alfa, aes(x = Damage_level, y = Evenness)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(),
               fill=rep(c("gray90","gray70"),2),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar",
                   color="black", width=0.2) 
e <- e + geom_jitter()+ facet_wrap(~Farm)+ theme_classic() 
e <- e +  xlab("") + ylab("Evenness")+ylim(c(0, max(Alfa$Evenness)*1.1))
e <- e + theme(axis.title =element_text(size=14,family = 'serif'))
e <- e +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,
                                             family ='serif'),
                  axis.text.y = element_text(face="bold",size=12, angle=0,family =
                                               'serif'))
e <- e + theme(strip.text.x = element_text(size = 14,family = 'serif'))
e <- e + geom_signif(test="kruskal.test", comparisons = list(c("Mild", "Severe")), map_signif_level = T)

```

*Figure 3.* Alfa diversity for the fungal communities of the avocado cv. Hass fruits with different severities of lenticel damage (Mild and Severe) collected from the La Escondida and El Sinai farms during the traviesa harvest of 2020. Bars and error bars represent mean and standard deviation for the alfa diversity index richness (A), Shannon (B), and Faith pd (C) alpha-diversity (n: 6). The points show the alfa diversity metrics the fruits and the asterisk denote statistic differences at the 95.0 % confidence level (*) according to the t test.

```{r severity and incidence plot}

 alpha_plot<-grid.arrange(s + labs(tag = 'A'),c + labs(tag = 'B'), e + labs(tag = 'C'),f + labs(tag = 'D'), ncol=2,nrow=2)
ggsave("Figure2_alfa_new.jpg",plot=alpha_plot,dpi = 300,width = 6,height = 6,units = c("in"))
```

## Beta diversity plots
Loading the ASVs table after CCS normalization and the phylogenetic tree for the Beta diversity annalysis

```{r uploading the CCS.biom and tree.nwk}

path<-"C:/Users/Dru/Desktop/Cartama_paper/microbial-ecology-avocado/Resultados_All"
biomfile <- import_biom(paste(path,"/css_eff.biom",sep=""))
biomfile_mapfile <- import_qiime_sample_data(paste(path,"/metadata_real.txt",sep=""))
tree <- read_tree(paste(path,"/tree.nwk",sep=""))

biomfile_mapfile_sinai<- subset(biomfile_mapfile,biomfile_mapfile$Farm1=="Sinai")
biomfile_mapfile_escondida<-subset(biomfile_mapfile,biomfile_mapfile$Farm1=="Escondida")

biomfile_merged <- merge_phyloseq(biomfile,biomfile_mapfile_escondida)
colnames(tax_table(biomfile_merged)) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
sam <-sample_data(biomfile_merged)
otu.table <- otu_table(biomfile_merged)
taxa <- tax_table(biomfile_merged)

otutax = phyloseq(otu.table, taxa)
merge2 = merge_phyloseq(otutax, sam, tree)
```

###CAP

```{r CAP}
agricultural.ord<- ordinate(merge2, "CAP", "wunifrac", ~Farm*Lenticel_disease_status)

agricultural_plot<- plot_ordination(merge2, agricultural.ord, type = "sites",
                                    color = "Farm",
                                    title = "", label = NULL)

CAP = agricultural_plot  +
  geom_point(aes(color =Farm,shape = Farm1), size = 3) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        rect = element_rect(fill = "white", colour = "black",size = 0.5, linetype = 1),
        axis.ticks = element_line(colour = "black"),
        axis.title.x = element_text(face="plain",family = "serif", colour="black", size=12),
        axis.title.y = element_text(face="plain",family = "serif", colour="black", size=12),
        axis.text.x  = element_text(face="plain",family = "serif", colour = "black", angle=360, vjust=0.5, size=10),
        axis.text.y  = element_text(face="plain",family = "serif", colour = "black", angle=360, vjust=0.5, size=10),
        panel.background = element_rect(fill = "white", colour = "black", linetype = 1, size = 0.5),
        panel.border = element_blank(), legend.title = element_blank(), legend.text=element_text(size=0.1,family = "serif"), plot.margin = margin(0.755, 1, 1, 0.3, "cm"))  +
  scale_colour_manual(values=c("gray85","gray55","gray85","gray55"),labels =
                        c("Escondida_Mild","Escondida_Severe",
                          "Sinai_Mild","Sinai_Severe")) + 
  stat_ellipse(type = "t", size=0.5) + labs(x = "CAP (23.7%)", y= "CAP (45.8%)") + theme(legend.position = "right",legend.text = element_text(size=10))
CAP
ggsave("Figure4_new.jpg",plot=CAP,dpi = 300,width = 6,height = 5,units = c("in"))

```

*Fig 4* Fungal communities’ structure for fruits of avocado cv. Hass fruits with mild and severe lenticel damages collected from the La Escondida and El Sinai farms during the traviesa harvest of 2020. Cumulative sum scaling (CSS) transformed reads were used to calculate weighted unifrac distance. Shown is the bidimensional plane of the canonical analysis of principal coordinates (CAP) constrained by the farm and the strenght of the damage (69.5 % of the overall variance; p-value: 0.001 acording to the permanova test). Each point represents a fruit's fungal community, the shapes show the farm (La Escondida (circles) and El Sinai (triangles)), and the colors the damage strength (Severe (dark gray) and Mild (light gray)) of the fruit. The ellipses are the 95 % confidence level ellipses.

```{r}

#agricultural.ord<- ordinate(merge2, "CAP", "wunifrac", ~Farm1*Lenticel_damage_status)

#PERMANOVA
perm_anova.agricultural.ord1 <- anova.cca(ordinate(merge2, "CAP", "wunifrac", ~Lenticel_disease_status))
#perm_anova.agricultural2.ord <- permutest(agricultural.ord)

# P-value
print(as.data.frame(perm_anova.agricultural.ord)["Pr(>F)"])
```

```{r PCoA and PCA plots}

po<-grid.arrange(PCoA+labs(tag = "A"),CAP+  geom_text(x=-0.65, y=0.69,label="B",size=4.5, col= "black") +coord_cartesian( clip = "off"),legend, ncol=2,nrow = 2, 
             layout_matrix = rbind(c(1,2), c(3,3)),
             widths = c(2.7, 2.7), heights = c(2.5, 0.2))
ggsave("C:/Users/Dru/Desktop/Cartama_paper/Figure4.jpg",plot=po,dpi = 300,width = 7,height = 6,units = c("in"))
```
## Taxonomy bat plots
Loading the ASVs table

```{r data barplot}
biomfile <- import_biom(paste(path,"/ASVs_table9_dada2R.biom",sep=""))
##biomfile <- import_biom("ASVs_table6_dada2R.biom")
#biomfile_mapfile <- import_qiime_sample_data(paste(path,"/metadata_real.txt",sep=""))
biomfile_merged <- merge_phyloseq(biomfile,biomfile_mapfile)
colnames(tax_table(biomfile_merged)) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
sample_data(biomfile_merged)

plot_colors <- c( "palegoldenrod","gray90", "plum4", "lightsteelblue","darkseagreen1",
                   "#CBD588", "#5F7FC7","#DA5724", "#8A7C64","gray30", "#508578", "#CD9BCD", "#AD6F3B",
                  "darkseagreen4","deepskyblue1","gray50" , "yellow2", "slateblue1",
                  "#673770","Khaki4", "#652926", "yellow3", "cadetblue3",
                   "lightsteelblue3", "honeydew3", "gray80","thistle4","yellow4",
                   "#5E738F", "#8A7C64","gray30", "lightpink3")
```

## Phylum barplot
```{r}
subset = biomfile_merged
rel.abun = transform_sample_counts(subset, function(x) x / sum(x))
glom <- tax_glom(rel.abun, taxrank = 'Phylum')
data_glom<- psmelt(glom) 
data_glom$Phylum <- as.character(data_glom$Phylum) 
data_glom$Phylum[data_glom$Abundance < 0.03] <- "< 3% abundance"
data_glom$Phylum[data_glom$Phylum == "unidentified"] <- "Not classified"
data_glom$Phylum[data_glom$Phylum == "undefined"] <-"Not classified"
data_glom$Lenticel_disease_status[data_glom$Lenticel_disease_status == "Necrotic"] <-"Severe"
data_glom$Lenticel_disease_status[data_glom$Lenticel_disease_status == "Healthy"] <-"Mild"

plot_colors <- c( "palegoldenrod","gray90", "plum4", "lightsteelblue","darkseagreen1")
Phylum <- ggplot(data_glom, aes(x = Lenticel_disease_status, y = Abundance/6, fill = reorder(Phylum, Abundance)))+
  geom_bar(stat = "identity") + facet_wrap(~Farm1) + scale_fill_manual(values =plot_colors) +
  ylab("Relative abundance") + xlab("") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        rect = element_rect(fill = "white",  colour = "black",  size = 0.5, linetype = 1),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.title.y = element_text(face="plain", colour="black", size=13,family="serif"),
        axis.text.x  = element_text(face="plain", colour = "black", angle=30, vjust=0.6, size=13,family="serif"),
        axis.text.y  = element_text(face="plain",colour = "black", angle=360, vjust=0.5, size=13,family="serif"),
        panel.background = element_rect(fill = "white", colour = "white", linetype = 1, size = 0.5),
        panel.border = element_blank()) + scale_y_continuous(labels = percent_format(accuracy = 1)) +
  theme(legend.title = element_blank(),legend.text = element_text(family="serif",size=12))+
  theme(strip.text.x = element_text(size = 13 , family = 'serif'))

print(Phylum)

df = data.frame(Treatment = data_glom$Farm, Phylum = data_glom$Phylum, Abundance = round(data_glom$Abundance,3)*100)
df = df[df$Phylum!="< 3% abundancia",]
df<-df[ order(df$Treatment,df$Phylum),]
mean<-tapply(df$Abundance, list(df$Phylum,df$Treatment),mean)
sd<-tapply(df$Abundance, list(df$Phylum,df$Treatment),sd)
sink("mean_phylum.txt")
print(mean)
sink()
sink("sd_phylum.txt")
print(sd)
sink()
```
## Class barplot
```{r}
subset = biomfile_merged
rel.abun = transform_sample_counts(subset, function(x) x / sum(x))
glom <- tax_glom(rel.abun, taxrank = 'Class')
data_glom<- psmelt(glom) 
data_glom$Class <- as.character(data_glom$Class) 
data_glom$Class[data_glom$Abundance < 0.03] <- "< 3% abundance"
data_glom$Class[data_glom$Class == "unidentified"] <- "Not classified"
data_glom$Class[data_glom$Class == "undefined"] <-"Not classified"

data_glom$Lenticel_disease_status[data_glom$Lenticel_disease_status == "Necrotic"] <-"Severe"
data_glom$Lenticel_disease_status[data_glom$Lenticel_disease_status == "Healthy"] <-"Mild"

plot_colors <- c( "palegoldenrod","gray40","thistle1", "lightsteelblue","darkseagreen1",
                   "#CBD588", "plum4", "#652926","Khaki4", "yellow3", "cadetblue3",
                   "lightsteelblue3", "honeydew3",

                  "gray80", "yellow2", "slateblue1", "#5F7FC7","#DA5724", "#8A7C64",
                  "gray30", "#508578", "#CD9BCD", "gray80","#AD6F3B",
                  "darkseagreen4","deepskyblue1","gray50" , "yellow2", "slateblue1",
                  "#673770","thistle4","yellow4",
                   "#5E738F", "#8A7C64","gray30", "lightpink3")

Class <- ggplot(data_glom, aes(x = Lenticel_disease_status, y = Abundance/6, fill = reorder(Class, Abundance)))+
  geom_bar(stat = "identity") + facet_wrap(~Farm1) + scale_fill_manual(values =plot_colors) +
  ylab("Relative abundance") + xlab("") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        rect = element_rect(fill = "white",  colour = "black",  size = 0.5, linetype = 1),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.title.y = element_text(face="plain", colour="black", size=13,family="serif"),
        axis.text.x  = element_text(face="plain", colour = "black", angle=30, vjust=0.6, size=13,family="serif"),
        axis.text.y  = element_text(face="plain",colour = "black", angle=360, vjust=0.5, size=13,family="serif"),
        panel.background = element_rect(fill = "white", colour = "white", linetype = 1, size = 0.5),
        panel.border = element_blank()) + scale_y_continuous(labels = percent_format(accuracy = 1)) +
  theme(legend.title = element_blank(),legend.text = element_text(family="serif",size=12))+
  theme(strip.text.x = element_text(size = 13 , family = 'serif'))

print(Class)

df = data.frame(Treatment = data_glom$Farm, Class = data_glom$Class, Abundance = round(data_glom$Abundance,3)*100)
df = df[df$Class!="< 3% abundancia",]
df<-df[ order(df$Treatment,df$Class),]
mean<-tapply(df$Abundance, list(df$Class,df$Treatment),mean)
sd<-tapply(df$Abundance, list(df$Class,df$Treatment),sd)
sink("mean_Class.txt")
print(mean)
sink()
sink("sd_Class.txt")
print(sd)
sink()
```
```{r}

Phylum <- Phylum + theme(legend.key.height= unit(0.2, 'cm'),
        legend.key.width= unit(0.2, 'cm'))

Class <- Class + theme(legend.key.height= unit(0.2, 'cm'),
        legend.key.width= unit(0.2, 'cm'))

po<-ggarrange(ggarrange(Phylum+labs(tag= "A"),Class+labs(tag= "B"), nrow= 2)) 
po
ggsave("Fig4.jpg",plot=po,dpi = 300,width = 6,height = 8,units = c("in"))

```

### diferential abundance analysis

loading the data for diferencial abundance annalysis-Escondida

```{r escondida diferential abundance, message=FALSE, warning=FALSE}

biomfile <- import_biom("ASVs_enrich_Escondida.biom")
biomfile_mapfile <- import_qiime_sample_data("metadata_real1.txt")
tree <- read_tree("tree.nwk")
biomfile_merged <- merge_phyloseq(biomfile,biomfile_mapfile)
colnames(tax_table(biomfile_merged)) <- c( "Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
sam <-sample_data(biomfile_merged)
otu.table <- otu_table(biomfile_merged)

otu.table<-otu.table[!(otu.table[,1]<0.005&otu.table[,2]<0.005),]
#over 0.5%
taxa <- tax_table(biomfile_merged)
otutax = phyloseq(otu.table, taxa)
GP.chl = merge_phyloseq(otutax, sam, tree)

```

```{r making a df}

labels_new<-c("ASV22_Not_classified", "ASV57_g_Erythrobasidium", "ASV31_c_Cystobasidiomycetes", 
              "ASV30_g_Cystobasidium", "ASV53_g_Golubevia", "ASV25_g_Meira", "ASV75_g_Vishniacozyma",
              "ASV11_g_Vishniacozyma", "ASV15_g_Vishniacozyma", "ASV24_g_Setophoma","ASV34_g_Bacidina",
              "ASV18_g_Bacidina", "ASV6_g_Hyphozyma",  "ASV4_g_Cladosporium",  "ASV1_Not_classified",
              "ASV2_Not_classified",  "ASV12_Not_classified", "ASV3_Not_classified", "ASV7_Not_classified",
              "ASV42_Not_classified", "ASV41_Not_classified", "ASV60_g_Neodevriesia", "ASV9_Not_classified",
              "ASV10_o_Chaetothyriales", "ASV13_Not_classified")

tree_new<-phy_tree(GP.chl)
tree_new$tip.label<-labels_new
otu_table_new<-otu_table(GP.chl)
row.names(otu_table_new)<-labels_new
taxa_new<-tax_table(GP.chl)
row.names(taxa_new)<-labels_new
sample_data_new<-sample_data(GP.chl)
otutax_new = phyloseq(otu_table_new, taxa_new)
merge_new = merge_phyloseq(otutax_new, sample_data_new, tree_new)

df<- as.data.frame(otu_table_new)
df$Ratio <- df$Escondida_Severo/df$Escondida_Leve
df$ASVs<- phy_tree(tree_new)$tip.label
df$Enriched_in[df$Ratio < 1]<- "Mildly damaged fruits"
df$Enriched_in[df$Ratio >= 1]<- "Severely damaged fruits"

df1 <- data.frame(
  ASVs = phy_tree(tree_new)$tip.label, Ratio = df$Ratio, Enriched_in = df$Enriched_in 
)

df_anotations <- data.frame(node = c(1,28,35), 
                            taxa = c("Not lassified", "Basidiomycetes", "Ascomycetes"))

df_anotations_class <- data.frame(node = c(41,23,49,38,13,10,30,32,33,1), 
                                  Class = c("Dothideomycetes", "Not classified", "Eurotiomycetes", "Lecanoromycetes", 
                                            "Leotiomycetes", "Dothideomycetes", "Cystobasidiomycetes","Exobasidiomycetes",
                                            "Tremellomycetes", "Not classified"))

df_anotations_class$Class <- factor(df_anotations_class$Class, 
                                    levels = c("Dothideomycetes", "Not classified", "Eurotiomycetes", 
                                               "Lecanoromycetes", "Leotiomycetes", "Cystobasidiomycetes",
                                               "Exobasidiomycetes","Tremellomycetes"))
#p + geom_text(aes(label=node), hjust=-.3)

```
## Escondida plot

*Fig 6.* Taxonomy and relative abundance of the amplicon sequence variants (ASVs) enriched (p-value < 0.05) in the fungal communities of avocado cv. Hass fruits with different severities of lenticel damage (Mild and Severe) collected from the La Escondida during the traviesa harvest of 2020. Shown are the taxonomic relation (A) and the relative abundances (B) of the ASVs group by color according to the clade (Supplementary Table 5).

```{r ploting the data}

p <- ggtree(tree_new) + geom_tiplab(align=T, size=3) + xlim(-.1, 3)+ hexpand(.01)+ labs(tag = "A")

p <- p + geom_hilight(data=df_anotations_class, aes(node=node, fill=Class),  alpha=0.2,  extend=3, type = "roundrect")
p<-p+geom_nodepoint(aes(subset = node == 35, x = x - .07),size = 3, colour = 'brown3', shape = 19)
p<-p+geom_nodepoint(aes(subset = node == 28, x = x - 0.12),size = 3, colour = 'deepskyblue3', shape = 19)
p<- p +   theme(legend.title = element_text(family="serif",size=12, face = "bold"),legend.text = element_text(family="serif",size=10))

p1 <- ggplot(df1, aes(x = ASVs, y = log10(Ratio))) + geom_col(aes(fill=Enriched_in)) + theme_classic() + coord_flip() + labs(tag = "B")
p1 <- p1 + scale_fill_manual("Enriched in:",values= alpha(c("Mildly damaged fruits" = "#00A087B2", "Severely damaged fruits" ="#7E6148B2"),0.4))
p1 <- p1 + ylab("log10(Abundance severe/Abundance mild)") + xlab("") + guides(y = "none") + theme(axis.text.y=element_blank(),axis.ticks.y=element_blank())
p1<- p1+  theme(axis.title.x = element_text(face="plain", colour="black", size=11,family="serif"),
        axis.text.x  = element_text(face="plain",colour = "black", angle=360, vjust=0.5, size=10,family="serif")) +
  theme(legend.title = element_text(family="serif",size=12, face = "bold"),legend.text = element_text(family="serif",size=10))

po <- p1 %>% insert_left(p)

#####

data <- data.frame(Xdata = rnorm(2), Ydata = rnorm(2),Phylum =c("Ascomycete","Basidiomycete"))

gplot <- ggplot(data, aes(Xdata, Ydata, color = Phylum)) + geom_point(size = 3) +scale_color_manual(values = c("Ascomycete" = "brown3", "Basidiomycete" = "deepskyblue3"))
gplot <- gplot +   theme(legend.title = element_text(family="serif",size=12, face = "bold"),legend.text = element_text(family="serif",size=10),legend.key = element_rect(colour = NA, fill = NA))

jpeg("Fig5.jpg",width = 8,height = 6,units = c("in"), res=300) 
po
leg=g_legend(gplot)
wc_table_vp <- viewport(x = 0.74, y = 0.20, just = c("left", "top"),height = 0.1, width = 0.2)
pushViewport(wc_table_vp)
grid.draw(leg)

dev.off()

```

loading the data for diferencial abundance annalysis-Sinai

```{r Sinai diferential abundance, message=FALSE, warning=FALSE, paged.print=FALSE}

# biomfile <- import_biom("ASVs_enrich_Sinai.biom")

#biomfile_mapfile <- import_qiime_sample_data("metadata_real1.txt")
#tree <- read_tree("tree.nwk")
biomfile_merged <- merge_phyloseq(biomfile,biomfile_mapfile)
colnames(tax_table(biomfile_merged)) <- c( "Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
sam <-sample_data(biomfile_merged)
otu.table <- otu_table(biomfile_merged)

otu.table<-otu.table[!(otu.table[,1]<0.005&otu.table[,2]<0.005),]
#over 0.5%
taxa <- tax_table(biomfile_merged)
otutax = phyloseq(otu.table, taxa)
GP.chl = merge_phyloseq(otutax, sam, tree)

```

```{r making a df sinai}

labels_new<-c("ASV37_Not_classified", "ASV96_Not_classified","ASV33_g_Symmetrospora","ASV20_g_Symmetrospora","ASV31_c_Cystobasidiomycetes",
              "ASV54_c_Cystobasidiomycetes","ASV11_g_Vishniacozyma","ASV65_g_Colletotrichum", "ASV40_g_Geastrumia", "ASV14_g_Pseudocercospora", 
              "ASV4_g_Cladosporium",  "ASV2_Not_classified",  "ASV3_Not_classified",  "ASV7_Not_classified",  "ASV70_Not_classified",
              "ASV10_o_Chaetothyriales", "ASV13Not_classified", "ASV26_g_Aureobasidium")

tree_new<-phy_tree(GP.chl)
tree_new$tip.label<-labels_new
otu_table_new<-otu_table(GP.chl)
row.names(otu_table_new)<-labels_new
taxa_new<-tax_table(GP.chl)
row.names(taxa_new)<-labels_new
sample_data_new<-sample_data(GP.chl)
otutax_new = phyloseq(otu_table_new, taxa_new)
merge_new = merge_phyloseq(otutax_new, sample_data_new, tree_new)

df<- as.data.frame(otu_table_new)
df$Ratio <- df$Sinai_Severo/df$Sinai_Leve
df$ASVs<- phy_tree(tree_new)$tip.label
df$Enriched_in[df$Ratio < 1]<- "Mildly damaged fruits"
df$Enriched_in[df$Ratio >= 1]<- "Severely damaged fruits"

df1 <- data.frame(
  ASVs = phy_tree(tree_new)$tip.label, Ratio = df$Ratio, Enriched_in = df$Enriched_in 
)

df_anotations <- data.frame(node = c(20,22,26), 
                            taxa = c("Not lassified", "Basidiomycetes", "Ascomycetes"))

df_anotations_class <- data.frame(node = c(20,7,23,
                                           8, 18,9, 35, 30), 
                                  Class = c("Not classified","Tremellomycetes", "Cystobasidiomycetes",
                                            "Sordariomycetes", "Dothideomycetes" ,"Dothideomycetes", 
                                            "Eurotiomycetes","Dothideomycetes"))

df_anotations_class$Class <- factor(df_anotations_class$Class, 
                                    levels = c("Dothideomycetes", "Eurotiomycetes","Sordariomycetes", 
                                               "Cystobasidiomycetes","Tremellomycetes","Not classified"))
#p + geom_text(aes(label=node), hjust=-.3)

```
*Fig 7.* Taxonomy and relative abundance of the amplicon sequence variants (ASVs) enriched (p- value < 0.05) in the fungal communities of avocado cv. Hass fruits with different severities of lenticel damage (Mild and Severe) collected from the La Sinai during the traviesa harvest of 2020. Shown are the taxonomic relation (A) and the relative abundances (B) of the ASVs group by color according to the clade (Supplementary Table 6).

```{r ploting the Sinai}

p <- ggtree(tree_new) + geom_tiplab(align=T, size=3) + xlim(-.1, 3)+ hexpand(.01)+ labs(tag = "A")

p <- p + geom_hilight(data=df_anotations_class, aes(node=node, fill=Class),  alpha=0.2,  extend=3, type = "roundrect")
p<-p+geom_nodepoint(aes(subset = node == 26, x = x - .13),size = 3, colour = 'brown3', shape = 19)
p<-p+geom_nodepoint(aes(subset = node == 22, x = x - 0.124),size = 3, colour = 'deepskyblue3', shape = 19)
p<- p +   theme(legend.title = element_text(family="serif",size=12, face = "bold"),legend.text = element_text(family="serif",size=10))

p1 <- ggplot(df1, aes(x = ASVs, y = log10(Ratio))) + geom_col(aes(fill=Enriched_in)) + theme_classic() + coord_flip() + labs(tag = "B")
p1 <- p1 + scale_fill_manual("Enriched in:",values= alpha(c("Mildly damaged fruits" = "#00A087B2", "Severely damaged fruits" ="#7E6148B2"),0.4))
p1 <- p1 + ylab("log10(Abundance severe/Abundance mild)") + xlab("") + guides(y = "none") + theme(axis.text.y=element_blank(),axis.ticks.y=element_blank())
p1<- p1+  theme(axis.title.x = element_text(face="plain", colour="black", size=11,family="serif"),
        axis.text.x  = element_text(face="plain",colour = "black", angle=360, vjust=0.5, size=10,family="serif")) +
  theme(legend.title = element_text(family="serif",size=12, face = "bold"),legend.text = element_text(family="serif",size=10))

po <- p1 %>% insert_left(p)

jpeg("Fig6.jpg",width = 8,height = 6,units = c("in"), res=300) 
po
leg=g_legend(gplot)
wc_table_vp <- viewport(x = 0.74, y = 0.20, just = c("left", "top"),height = 0.1, width = 0.2)
pushViewport(wc_table_vp)
grid.draw(leg)

dev.off()

ggsave("Fig6.jpg",plot=po,dpi = 300,width = 8,height = 6,units = c("in"))
```

###PCoA

```{r PCoA}

example.ord <- ordinate(merge2, "PCoA", "Unifrac", weighted=TRUE)

PCoA<- plot_ordination(merge2, example.ord, type = "sites", color = "Farm", label = NULL)
PCoA = PCoA  + 
  geom_point(aes(color =Farm,shape = Farm1), size = 3) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        rect = element_rect(fill = "white", colour = "black",size = 0.5, linetype = 1),
        axis.ticks = element_line(colour = "black"),
        axis.title.x = element_text(face="plain",family = "serif", colour="black", size=12),
        axis.title.y = element_text(face="plain",family = "serif", colour="black", size=12),
        axis.text.x  = element_text(face="plain",family = "serif", colour = "black", angle=360, vjust=0.5, size=10),
        axis.text.y  = element_text(face="plain",family = "serif", colour = "black", angle=360, vjust=0.5, size=10),
        panel.background = element_rect(fill = "white", colour = "black", linetype = 1, size = 0.5),
        panel.border = element_blank(), legend.title = element_blank(), legend.text=element_text(size=10,family = "serif"), plot.margin = margin(1, 0.5, 1, 1, "cm"))  +
  scale_colour_manual(values=c("gray85","gray55","gray85","gray55"),labels = c("Escondida_Mild","Escondida_Severe", "Sinai_Mild","Sinai_Severe"))+  
  stat_ellipse(type = "t", size=0.5) + theme(legend.position = "bottom") +
  labs(x = "PCoA (60.4%)", y= "PCoA (12.2%)")+ guides(colour = guide_legend(order = 2), 
              shape = guide_legend(order = 1))
legend <- get_legend(PCoA)
PCoA <- PCoA + theme(legend.position="none")

```