### 2023-04-01 R/Principal_lentcelosisa.Rmd

---
title: "Principal_analysis_lenticelosis"
author: "dru"
output: html_document
date: "2022-08-16"
---

## Preparing the environment

Loading required Packages

```{r Packages, echo=TRUE, message=FALSE, warning=FALSE}

library(glmmTMB)
library(bbmle)
library(DHARMa)
library(readxl)
library(tidyverse)

```

## Loading and inspecting the data

Loading and inspecting the file 'Lenticel_Damage_Summary_All_Harvest.xlsx'. This file has two-years of lenticel damage measurements recorded in two avocado farms across four harvest.

```{r echo=TRUE, message=FALSE, warning=FALSE}
Lenticel_file <- read_excel("C:/Users/Dru/Desktop/Cartama_paper/files/Lenticel_Damage_Summary_All_Harvest.xlsx")
glimpse(Lenticel_file)
```

Checking for NA, miss speels and anomalies in the data:

```{r inspecting}

table(Lenticel_file$Cosecha)
table(Lenticel_file$UP)

f <-subset(Lenticel_file,Lenticel_file$UP=="Escondida")
f1 <- subset(Lenticel_file,Lenticel_file$UP=="Sinai")
table(f$Lote)
table(f$Arbol)
table(f1$Lote)
table(f1$Arbol)
summary(Lenticel_file$Severidad_0)
summary(Lenticel_file$Severidad_1)
summary(Lenticel_file$Incidencia_0)
summary(Lenticel_file$Incidencia_1)
```

Problems:

* Some tree numbers are shared between La Escondida y El Sinai, and this redundancy can result in problems during the analysis.
* Lenticel damage columns have several NA. Fruits having NA the severidad_0 or incidence_0 were removed from all the analyses, and those having NA in severidad_1 or incidence_1 only from the analysis evaluating disease progresion

## Cleaning and preparing the data
### cleaning actions:

* The columns cosecha, Año, Severidad_pedunculo, arbol and the last three empty columns were removed as they are not used during the analysis
* The 'p' simbol in the column Cosecha is changed for 'm' to especify the main harvest.
* Generated a column Tree which has the name of the tree indicated by the farm and tree number (i.e., Sinai15) to avoid problems during the analysis

```{r cleaning}
Lenticel_file$...13<-NULL
Lenticel_file$...14<-NULL
Lenticel_file$...15<-NULL
Lenticel_file$Año<-NULL
Lenticel_file$cosecha<-NULL
Lenticel_file$Severidad_pedunculo<-NULL
Lenticel_file[Lenticel_file=="2019_p"]<-"2019_m"
Lenticel_file[Lenticel_file=="2020_p"]<-"2020_m"
Lenticel_file$Tree<-paste(Lenticel_file$UP,Lenticel_file$Arbol,sep="")
Lenticel_file$Arbol<-NULL

df <- Lenticel_file %>% drop_na(Severidad_0 , Incidencia_0)
colnames(df)[which(names(df) == "Cosecha")] <- "Harvest"
colnames(df)[which(names(df) == "UP")] <- "Farm"
colnames(df)[which(names(df) == "Lote")] <- "Plot"
colnames(df)[which(names(df) == "Fruta")] <- "Fruit"
colnames(df)[which(names(df) == "Severidad_0")] <- "Sev_0dph"
colnames(df)[which(names(df) == "Severidad_1")] <- "Sev_21dph"
colnames(df)[which(names(df) == "Incidencia_0")] <- "Inc_0dph"
colnames(df)[which(names(df) == "Incidencia_1")] <- "Inc_21dph"
df$Inc_ratio <- df$Inc_21dph-df$Inc_0dph
df$Sev_ratio <- df$Sev_21dph-df$Sev_0dph
glimpse(df)
summary(df$Sev_0dph)

#Datos_mat<-subset(Datos_mat,Datos_mat$Cosecha!="2020_t")

plot<-factor(df$Plot)
tree<-factor(df$Tree)
farm<-factor(df$Farm)
fpt<-farm:plot:tree
fp<-farm:plot
pt<- plot:tree
year<-factor(df$Harvest)
fruta<-factor(df$Fruit)
```
# 0dph annalysis for the severity and Incidence
## Building the severity models
We are using zero-inflated generalized linear mixed models for assessing differences in severity between farms and harvests. We are using glmmTMB function of the glmmTMB package with the beta family().
https://cran.r-project.org/web/packages/glmmTMB/vignettes/glmmTMB.pdf

```{r}

y<-df$Sev_0dph/100

modelo <- glmmTMB(y~farm*year +  (1|fp) + (1|fpt),ziformula=~1,family=beta_family())

modelo1 <- glmmTMB(y~farm*year + (1|farm) + (1|fp) + (1|fpt),ziformula=~1,family=beta_family())
#modelo1 <- glmmTMB(y~farm+year + (1|farm) + (1|fp) + (1|fpt),ziformula=~1,family=beta_family())
#modelo2 <- glmmTMB(y~year + (1|farm) + (1|fp) + (1|fpt),ziformula=~1,family=beta_family())
#modelo3 <- glmmTMB(y~farm + (1|farm) + (1|fp) + (1|fpt),ziformula=~1,family=beta_family())

```
## Testing the model
Testing the good-of-fit of the model with the DHARMA library
https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html

We used AIC model selection to distinguish among a set of possible models describing the relationship between farm and harvest and severity. The best-fit model, carrying 100% of the cumulative model weight, included every parameter with the interaction effects.
https://www.scribbr.com/statistics/akaike-information-criterion/#:~:text=The%20Akaike%20information%20criterion%20(AIC,best%20fit%20for%20the%20data.

```{r}

#AICtab(modelo,modelo1,modelo2,modelo3, base=TRUE,weights = TRUE)

residual_simulation<-simulateResiduals(modelo)
plot(residual_simulation)
residual_simulation_years <-recalculateResiduals(residual_simulation, group = df$Tree)
plot(residual_simulation_years)
testDispersion(residual_simulation)
testZeroInflation(residual_simulation)
testCategorical(residual_simulation, catPred = df$Tree)
par(mfrow = c(1,2))
plotResiduals(residual_simulation, farm)
plotResiduals(residual_simulation, year)

sink("Sev_0h.txt")
print(summary(modelo))
sink()
```

## Building the incidence models
We are using zero-inflated generalized linear mixed models for assessing differences between farms and years.  we are using glmmTMB function of the glmmTMB package with the negative binomial family.
https://cran.r-project.org/web/packages/glmmTMB/vignettes/glmmTMB.pdf

```{r}

y<-round(df$Inc_0dph)

modelo <- glmmTMB(y~farm*year + (1|farm) + (1|fp) + (1|fpt),ziformula=~1,family=nbinom2)
modelo1 <- glmmTMB(y~farm+year + (1|farm) + (1|fp) + (1|fpt),ziformula=~1,family=nbinom2)
modelo2 <- glmmTMB(y~year + (1|farm) + (1|fp) + (1|fpt),ziformula=~1,family=nbinom2)
modelo3 <- glmmTMB(y~farm + (1|farm) + (1|fp) + (1|fpt),ziformula=~1,family=nbinom2)

```
## Testing the model
Testing the model with the DHARMA library
https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html
We selected modelo1 as its better fitted and the AIC penalty is low. We are having issues with no homogen variance between harvest. We also have a small deviation from the expected distributions of the residualas

```{r}

AICtab(modelo,modelo1,modelo2,modelo3, base=TRUE)

residual_simulation<-simulateResiduals(modelo)
plot(residual_simulation)
residual_simulation_years <-recalculateResiduals(residual_simulation, group = df$Tree)
plot(residual_simulation_years)
testDispersion(residual_simulation)
testZeroInflation(residual_simulation)
#testCategorical(residual_simulation, catPred = df$Tree)
par(mfrow = c(1,2))
plotResiduals(residual_simulation, farm)
plotResiduals(residual_simulation, year)

sink("In_0h.txt")
print(summary(modelo))
sink()
```
# 21dph annalysis for the severity and Incidence
## Building the severity models
We are using zero-inflated generalized linear mixed models for assessing differences in severity between time points (0dph and 21dph) and harvests. These evlauations were mede for each farm separately using glmmTMB function of the glmmTMB package with the beta family().
https://cran.r-project.org/web/packages/glmmTMB/vignettes/glmmTMB.pdf

```{r}
dfe<-na.omit(subset(df,df$Farm=="Escondida"))
dfs<-na.omit(subset(df,df$Farm=="Sinai"))
df1<-na.omit(df)
y<-c(df1$Sev_0dph,df1$Sev_21dph)/100
ye<-c(dfe$Sev_0dph,dfe$Sev_21dph)/100
ys<-c(dfs$Sev_0dph,dfs$Sev_21dph)/100

time<-factor(c(rep("0dpc",length(y)/2),rep("21dpc",length(y)/2)))
timee<-factor(c(rep("0dpc",length(ye)/2),rep("21dpc",length(ye)/2)))
times<-factor(c(rep("0dpc",length(ys)/2),rep("21dpc",length(ys)/2)))

farm<-c(df1$Farm,df1$Farm)
f<-factor(c(df1$Fruit,df1$Fruit))
h<-factor(c(df1$Harvest,df1$Harvest))

fe<-factor(c(dfe$Fruit,dfe$Fruit))
he<-factor(c(dfe$Harvest,dfe$Harvest))
fs<-factor(c(dfs$Fruit,dfs$Fruit))
hs<-factor(c(dfs$Harvest,dfs$Harvest))

modelo <- glmmTMB(y ~ time * h *farm + (1|f)  ,ziformula= ~ 1,family=beta_family())
modelos <- glmmTMB(ys ~ times * hs + (1|fs) ,ziformula= ~ 1,family=beta_family())
modeloe <- glmmTMB(ye ~ timee * he + (1|fe) ,ziformula= ~ he + timee,family=beta_family())
```
## Testing the model
Testing the model with the DHARMA library
https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html
We selected modelos and modeloe

```{r}
#modelo <- glmmTMB(y ~ time * h *farm + (1|f)  ,ziformula= ~ 1,family=beta_family())
#modelo1 <-glmmTMB(y ~ time * h +farm + (1|f)  ,ziformula= ~ 1,family=beta_family()) 
#modelo2 <- glmmTMB(y ~ time + h *farm + (1|f)  ,ziformula= ~ 1,family=beta_family())
#modelo3 <- glmmTMB(y ~ time + h +farm + (1|f)  ,ziformula= ~ 1,family=beta_family())

AICtab(modelo,modelo1,modelo2,modelo3, base=TRUE,weights = TRUE)

residual_simulation<-simulateResiduals(modelo)
plot(residual_simulation)
testDispersion(residual_simulation)
testZeroInflation(residual_simulation)
par(mfrow = c(1,2))
plotResiduals(residual_simulation, h)

```
```{r}
dfe<-na.omit(subset(df,df$Farm=="Escondida"))
dfs<-na.omit(subset(df,df$Farm=="Sinai"))
df1<-na.omit(df)
y<-round(c(df1$Sev_0dph,df1$Sev_21dph))
ye<-round(c(dfe$Inc_0dph,dfe$Inc_21dph))
ys<-round(c(dfs$Inc_0dph,dfs$Inc_21dph))

time<-factor(c(rep("0dpc",length(y)/2),rep("21dpc",length(y)/2)))
timee<-factor(c(rep("0dpc",length(ye)/2),rep("21dpc",length(ye)/2)))
times<-factor(c(rep("0dpc",length(ys)/2),rep("21dpc",length(ys)/2)))

farm<-c(df1$Farm,df1$Farm)
f<-factor(c(df1$Fruit,df1$Fruit))
h<-factor(c(df1$Harvest,df1$Harvest))

fe<-factor(c(dfe$Fruit,dfe$Fruit))
he<-factor(c(dfe$Harvest,dfe$Harvest))
fs<-factor(c(dfs$Fruit,dfs$Fruit))
hs<-factor(c(dfs$Harvest,dfs$Harvest))

#modelo <- glmmTMB(y ~ time * h *farm + (1|f)  ,ziformula= ~ 1,family=beta_family())
modelo <- glmmTMB(y ~ time * h *farm + (1|f)  ,ziformula= ~ 1,family=nbinom2)
modelo1 <-glmmTMB(y ~ time * h +farm + (1|f)  ,ziformula= ~ 1,family=nbinom2) 
modelo2 <- glmmTMB(y ~ time + h *farm + (1|f)  ,ziformula= ~ 1,family=nbinom2)
modelo3 <- glmmTMB(y ~ time + h +farm + (1|f)  ,ziformula= ~ 1,family=nbinom2)

#modeloe <- glmmTMB(ye ~ timee * he + (1|fe) ,ziformula= ~ 1 + timee,family=nbinom2)
```
## Testing the model
Testing the model with the DHARMA library
https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html
We selected modelos and modeloe

```{r}
AICtab(modelo,modelo1,modelo2,modelo3, base=TRUE)
#modelos <- glmmTMB(ys ~ times * hs + (1|fs) ,ziformula= ~ 1,family=nbinom2)

residual_simulation<-simulateResiduals(modelo)
plot(residual_simulation)
testDispersion(residual_simulation)
testZeroInflation(residual_simulation)
plotResiduals(residual_simulation, h)

sink("In_0h_21h.txt")
print(summary(modelo))
sink()

```

```{r}

df1<-na.omit(df)

mat1<-data.frame(Farm =c(df1$Farm,df1$Farm), Time = factor(c(rep("0dpc",length(y)/2),rep("21dpc",length(y)/2))), 
                 Harvest = factor(c(df1$Harvest,df1$Harvest)), Severity = c(df1$Sev_0dph,df1$Sev_21dph)/100,
                 Incidence = round(c(df1$Inc_0dph,df1$Inc_0dph)), Fruit = c(df1$Fruit,df1$Fruit))

mat2 <-subset(mat1, mat1$Farm== "Sinai" & mat1$Harvest == "2020_m")
modelo <- glmmTMB(data=mat2, Severity ~ Time + (1|Fruit)  ,ziformula= ~ 1,family=beta_family())
residual_simulation<-simulateResiduals(modelo)
plot(residual_simulation)
testDispersion(residual_simulation)
summary(modelo)
modelo1 <- glmmTMB(data=mat2, Incidence ~ Time + (1|Fruit)  ,ziformula= ~ 1,family=poisson())
#modelo1 <- glmmTMB(data=mat2, Incidence ~ Time + (1|Fruit)  ,ziformula= ~ 1,family=nbinom2())
residual_simulation<-simulateResiduals(modelo1)
plot(residual_simulation)
testDispersion(residual_simulation)
summary(modelo1)

```